---
title: "R Markdown Presentation & Plotly"
author: "Nikhil Agrawal"
date: "2/19/2023"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	tidy = TRUE,
	tidy.opts = list(width.cutoff = 60)
)
```


## Project Description
Create a web page presentation using R Markdown that features a plot created with Plotly. 
Host your webpage on either GitHub Pages, RPubs, or NeoCities. 
Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly

```{r Date}
date()
```

## Plotly - Interactive Plot 1: Time-Series Chart

```{r echo=FALSE}
library(datasets)
library(plotly)
data(uspop) ## Load the data set that gives the population of the United States in millions for the period 1790-1970
## Plotly Command
plot_ly(x=~time(uspop),y=~uspop,type="scatter",mode="lines") %>% layout(title = "U.S. Population in millions for the period 1790-1970", xaxis = list(title = "Year"),yaxis = list(title = "U.S. Population (millions)"))
```


## Cars Summary

```{r cars}
summary(cars)
```

## Including Plots

```{r pressure, echo=FALSE}
plot(pressure)
```


## Thank you